Approximation as Differentiation

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Construct biproducts from coproducts on a cmon-category.

\noindent Since $(X, \meet, \top)$ is a commutative monoid, and thus a $\CMon$-enriched poset category, for
any $X, Y$ there is a zero meet-semilattice homomorphism $0_{X,Y}: X \to \top_Y$ and addition of homomorphisms
$f + g: X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join,
\bot)$. Now define the category $\LatGal$ where:
\begin{itemize}
\item objects $X, Y$ are the bounded lattices;
\item morphisms $f \vdash g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
$y \in Y$.
\end{itemize}

@@

$\LatGal$ is itself $\CMon$-enriched. For any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative monoid,
where:
\begin{itemize}
\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \vdash 0_{Y,X}: X \to Y$
\item $(f \vdash g) + (f' \vdash g') = (f + f') \vdash (g + g'): X \to Y$.
\end{itemize}

\noindent and biadditivity of composition arising component-wise. Moreover $\LatGal$ has binary products, with
the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts:

>>>>

\noindent Since $(X, \meet, \top)$ is a commutative monoid, and thus a $\CMon$-enriched poset category, for
any $X, Y$ there is a zero meet-semilattice homomorphism $0_{X,Y}: X \to \top_Y$ and addition of homomorphisms
$f + g: X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join,
\bot)$. Now define the category $\LatGal$ where:
\begin{itemize}
\item objects $X, Y$ are the bounded lattices;
\item morphisms $f \dashv g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
$y \in Y$.
\end{itemize}

@@

$\LatGal$ is itself $\CMon$-enriched. For any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative monoid,
where:
\begin{itemize}
\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \dashv 0_{Y,X}: X \to Y$
\item $(f \dashv g) + (f' \dashv g') = (f + f') \dashv (g + g'): X \to Y$.
\end{itemize}

\noindent and biadditivity of composition arises component-wise. Moreover $\LatGal$ has binary products, with
the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts
(\propref{biproduct:from-product-or-coproduct}).
